# Data Wrangling
In this chapter, I take a raw .CSV file and use data wrangling techniques to prepare
it for future analysis and graphing. 

## Load required R packages and raw data

I first need to clean up the data from the greenhouse experiment and get it into 
a logical order. Let's begin!
```{r Load packages}
# Load required packages
library(tidyverse)
```

```{r Load data}
# Load our data
setwd("../Raw_Data")
Germ <- read.csv("Germination.csv", header=T)
head(Germ)
# Drop columns we don't need
Germ <- Germ[ , -c(7, 8, 9)]
```

Tidyverse can be a clearer way to drop columns. For example, we could have written:

- Germ <- Germ %>% 
    select(-c(Observer:Checked_by)) 
and gotten the same result. 

## Changing column class
We need to ensure R is reading our data how we would like. Typically, this requires
changing the column class by specifying what format our data is in. But to do 
that - we need to know what class our data currently is!

```{r class}
# View default class
sapply(Germ, class)
```

We would like Species, Column, Row, Salinity, and Species_Tmt_Pos to be classed
as factors. Additionally, we want our Date column to be classed as a date.

```{r Change column class}
# A slow way to change column class is to call each column directly, like below
Germ$Species <- as.factor(Germ$Species)
class(Germ$Species)

# Here is a faster way to change multiple column classes
# Create a vector containing all of our future factor column names
factor_cols <- c("Species", "Column", "Row", "Salinity", "Species_Tmt_Pos")
# Use baseR functions to change all of these columns in one line of code
Germ[factor_cols]<-lapply(Germ[factor_cols], as.factor)
```

Dates are a special kind of class in R. We can use library(lubridate) to set this
up.
```{r dates}
# View how our dates are written
head(Germ, 2)
# These are written in mm dd yy format, so we can use the lubridate function mdy()
# to change the class
Germ$Date <- lubridate::mdy(Germ$Date)
```

Great! We should have our classes all in the right order now!
Let's check
```{r check class again}
sapply(Germ, class)
```

## Explore the data frame
Now we take a birds eye view of our data frame and check for typos, reasonable
values, and more.
```{r exploring}
# How many rows is our data frame?
length(Germ) 

# Do we have any NAs? and if so - where are they located in our dataframe?
which(is.na(Germ))

# View our min and max emergence to check for reasonable values
# We could do this with baseR
min(Germ$Emergence)
max(Germ$Emergence)
# or with Tidyverse
Germ %>% 
  select(Emergence) %>% 
  max()

# We can also check our factor levels to see if we have any typos, or unexpected
# values, etc..
levels(Germ$Species)
levels(Germ$Salinity)
levels(Germ$Column)
levels(Germ$Row)

# Alternatively, we can use a Lapply function to view levels of each factor column
# at once!
lapply(Germ, levels)
```

## Adding a column
Our dataframe includes monitoring date, but it would be more intuitive if we had
a column that specified experiment day rather than date. Let's add one.
 
```{r add experiment day column}
# View how many unique dates we recorded emergence
unique(Germ$Date)

# Alternatively in tidyverse...
Germ %>% 
  select(Date) %>% 
  distinct()

# Add a column where 01-30 is day 3 of the experiment, and then continued..
Germ <- Germ %>%
  mutate(Day = dplyr::case_when(
                        Date == "2023-01-30" ~ 3,
                        Date == "2023-02-02" ~ 6,
                        Date == "2023-02-05" ~ 9,
                        Date == "2023-02-08" ~ 12,
                        Date == "2023-02-11" ~ 15,
                        Date == "2023-02-14" ~ 18,
                        Date == "2023-02-17" ~ 21,
                        Date == "2023-02-20" ~ 24,
                        Date == "2023-02-23" ~ 27,
                        Date == "2023-02-27" ~ 31))

```
## Removing (filtering) data 
Although we evaluated 4 species in this experiment, we need to remove Distichlis spicata,
(or DISP), due to low emergence in all treatments. This species typically performs
well in our experiments, so we assume that its low emergence was due to some 
issue with this particular seed lot.
```{r remove DISP}
Germ <- Germ %>% 
  filter(Species != "DISP")

```

## Changing incorrect data
Last, we initially recorded 2 emergence counts for Salicornia rubra (SARU) at the
40 PPT salinity level. At a later date, we decided that the "emergence" was more 
likely the radicle unfurling with imbibition, but was not necessarily true emergence.
To address this we will change the counts of emergence to 0 at these levels
```{r change values, echo = TRUE}
Germ$Emergence[Germ$Species == "SARU" & Germ$Salinity == "40" & 
                       Germ$Emergence != "0"] <- 0
```

## Save your data object for future use
Whew! Now that we've gone through all this effort we can save the cleaned data 
object and simply load it for future coding use.

```{r save object, echo = TRUE, eval = FALSE}
getwd()
setwd("../Processed_Data")
save(Germ, file="DF_Germination.Rdata")
list.files()
```


