# Creating Visuals with our dataset
In this chapter, I will create a few graphs showing raw data results 
from the experiment.

## Load our previously cleaned dataset

```{r load processed data}
getwd()
setwd("../Processed_Data")
list.files()
load("DF_Germination.RData")
```

We also want to load any required packages. We could call these individually,
or we could simply load tidyverse
```{r load tidy}
library(tidyverse)
```

## Prepare dataset to create a graph showing cumulative germination counts
```{r view dataframe}
head(Germ)
```
Here, we can see that we have emergence counts for each species and salinity level,
but these are not cumulative counts. We need to do several things. 1) create a 
cumulative emergence column 2) average across the species and salinity interaction, 
and 3) create a standard error of these values.

```{r cumsum}
# Create a standard error function
std.error <- function(x) sd(x)/sqrt(length(x))

Germ <- Germ %>% 
  group_by(Species_Tmt_Pos) %>% 
  mutate(Total_Germination = cumsum(Emergence)) %>% 
  ungroup() %>% 
  mutate(Salinity = factor(Salinity, levels = c("Fresh", "25", "40"))) %>% 
  group_by(Species, Salinity, Day) %>% 
  summarize(SE = std.error(Total_Germination),
                   Total_Germination = mean(Total_Germination)) %>% 
  ungroup() %>% 
  as_tibble()
```

## Cumulative Germination Graph
```{r cumemerg graph}
vec <- c("darkblue", "blue", "lightblue")

ggplot(Germ, 
       aes(x = Day, y = Total_Germination, color = Species)) +
  geom_point(size = 4) +
  geom_line() +
  geom_linerange(ymin = Germ$Total_Germination - Germ$SE,
                 ymax = Germ$Total_Germination + Germ$SE) +
  labs(title = "Cumulative Germination",
       x = "Monitoring Day",
       y = "Mean Cumulative Germination ± SE") +
  theme(axis.title = element_text(size = 18),
        plot.title = element_text(size = 22, hjust = 0.5),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        ) +
  scale_color_manual(values = vec) +
  ylim(0, 100) + # Set 100 upper limit since that is what we seeded
  facet_wrap(~ Salinity) +
  theme_bw() 

```
Above we created a graph that was faceted by salinity. Now, we would like to facet
by species to view species-specific tolerances to salinity.

## Cumulative Germination Graph
```{r species facet}

vec <- c("blue", "yellow", "red")

ggplot(Germ, 
       aes(x = Day, y = Total_Germination, color = Salinity)) +
  geom_point(size = 4) +
  geom_line() +
  geom_linerange(ymin = Germ$Total_Germination - Germ$SE,
                 ymax = Germ$Total_Germination + Germ$SE) +
  labs(title = "Cumulative Germination",
       x = "Monitoring Day",
       y = "Mean Cumulative Germination ± SE") +
  theme(axis.title = element_text(size = 18),
        plot.title = element_text(size = 22, hjust = 0.5),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        ) +
  scale_color_manual(values = vec) +
  ylim(0, 100) + # Set 100 upper limit since that is what we seeded
  facet_wrap(~ Species) +
  theme_bw() 

```



  

