# Analyze data
We've now seen what our raw data looks like, and we have an idea that salinity
had a significant effect on species germination. However, we would like to show
that effect using a generalized linear mixed model.

## Load data.
```{r load data ch4, echo = FALSE}
setwd("../Processed_Data")
load("DF_Germination.Rdata")
```

Load packages too.
```{r load packages ch4}
library(tidyverse) #gives ggplot2 and piping
library(glmmTMB)
library(car)
library(DHARMa)
library(emmeans)
library(blme)
```

We will analyze our data using total germination counts, and will likely use a
Poisson distribution. To do that, we need to create a new column for total counts.
We also want to add a new column which accounts for the location (or block) in the
greenhouse. We will call this spot the 'Whole_Plot'

```{r total germ col}
# Whole Plot
Germ$Whole_Plot <- as.factor(paste(Germ$Column, Germ$Row))

# Total Emergence
Germ <- Germ %>%
  dplyr::group_by(Species, Salinity, Whole_Plot) %>%
  dplyr::summarise(Total_Emergence = sum(Emergence)) %>%
  dplyr::ungroup()

```
## Remove 40 PPT
If you recall, the 40 PPT had zero germination counts. We recognize that this is
a species tolerance limit for ALL of our species in this experiment. We won't gain
any new information by including this in our analysis, so we remove it here.
```{r remove 40 PPT}
Germ <- Germ %>%
  filter(Salinity %in% c("Fresh", "25")) %>% 
  mutate(Salinity = factor(Salinity, levels = c("Fresh", "25")))
```
We can check that we successfully removed 40 as a salinity factor.
```{r check factor 40}
levels(Germ$Salinity)
```

## Poisson model
The first stab at model creation.
```{r first mod}
G_Mod <- glmmTMB(Total_Emergence ~ Species + Salinity + Species:Salinity
            +  (1 | Whole_Plot), 
            family = poisson, # count data
            data = Germ)
```

## Poisson model diagnostic plots
Let's see how this model fits by looking at residual plots and DHARMa diagnostics.
```{r DHARMa diagnostics basic, echo=FALSE}
G_Mod.resid <- simulateResiduals(G_Mod, plot=T)
plotResiduals(G_Mod.resid, form = Germ$Salinity)
```

Our model fit could be improved. The second plot shows us that our mean very likely
does not equal our variance (as is assumed by a Poisson distribution). To address this,
we can swap to a negative binomial distribution and see if the diagnostics improve.

## Negative binomial model
```{r neg binom model}
G_Mod <- glmmTMB(Total_Emergence ~ Species + Salinity + Species:Salinity
                  +  (1 | Whole_Plot),
                  family = nbinom1(), # quasi poisson
                  data = Germ)
```
Note that our model statement is exactly the same, with the exception of the 
family call. Let's look at those diagnostic plots.

```{r diagnostic plots nbinom, echo = FALSE}
G_Mod.resid <- simulateResiduals(G_Mod, plot=T)
```

This looks better! Now I will explore the model further.

```{r explore nbinom model, echo = FALSE}
plotResiduals(G_Mod.resid, form = Germ$Salinity)
```

The spread of residuals is improved here. We can also look at the salinity effect
on our model estimates.

```{r other nbinom, echo = FALSE}
emmip(G_Mod, Species ~ Salinity, CIs = F)

```
The Salicornia rubra here has a highly negative prediction from Fresh to 25.
This poses an issue when interpreting our data. It likely resulted from having 
zero germination at the 25 PPT salinity for SARU (as seen below).

```{r SARU zero, echo = FALSE}
Germ %>% 
  filter(Salinity == "25" & Species == "SARU")
```

We had 4 replicates, and you can see that each replicate had zero seeds emerge.
Poisson and negative binomial distributions struggle with zero counts, because it 
has (1) no variance and (2) it exists right at the edge of the distributions range 
(0, infinity). We need to address this by adding a fixed effects prior to our
model. Our model interpretation is meaningless as is, because the 0's can impact
other model estimates.

## Fixed effects prior
Below we switch to using bglmer() rather than glmmTMB() because it allows us to add
fixed effects priors. These are added to our model estimates, in the order of our
model summary. To remind ourselves how many fixed effects we have, we can look at our
summary.

```{r summ FE, echo = FALSE}
summary(G_Mod)
```
There are 6 fixed effects to account for. Also note that the estimate for SARU 
and its interaction with 25 PPT has a really large standard error (1.183e04). 
That number should change if we successfully adjust away from 0.

```{r fixed effects model}
G_Mod <- bglmer(Total_Emergence ~ Species * Salinity
                  +  (1 | Whole_Plot),
                  family = poisson(), # back to Poisson for now
                  fixef.prior = normal(sd = c(3, 3, 3, 3, 3, 3)), 
                  data = Germ)
```
Priors of 3 represent a reasonable value (2.5 is default). The larger number 
you have the less of an effect it makes on your data. 

Let's see if the standard error was adjusted.
```{r FE summ, echo = FALSE}
summary(G_Mod)
```
Now the standard error is 1.58, which is a lot more reasonable! However, we may
not wish to penalize all of our fixed effects if SARU poses the only issue. In
this case, we could use large numbers on effects where we want weak penalization,
and use smaller numbers (e.g., 2.5) for our problem child. One more note that PHAU 
(Phragmites australis) also had low germination at 25 PPT. To account for this, 
I will penalize PHAU:25 as well.

## Weaker fixed effects priors

```{r weak FE}
G_Mod <- bglmer(Total_Emergence ~ Species * Salinity
                 +  (1 | Whole_Plot),
                 family = poisson(),
                 fixef.prior = normal(c(1e4, 1e4, 1e4, 1e4, 2.5, 2.5)), 
                 data = Germ)
summary(G_Mod)
```

We can explore this model to assess model fit.
```{r explore model, echo = FALSE}
G_Mod.resid <- simulateResiduals(G_Mod, plot=T)
plotResiduals(G_Mod.resid, form = Germ$Salinity)
```
  This model fit could be improved. Earlier, we fixed these issues by moving to a
nbinom() distribution. Unfortunately, that is not possible when using fixed effects
priors. 

However, let's take a step back and look at the raw data.
```{r graphs again, echo = FALSE}
# Create a standard error function
std.error <- function(x) sd(x)/sqrt(length(x))
vec <- c("darkblue", "blue", "lightblue")

setwd("../Processed_Data")
load("DF_Germination.RData")

Germ %>% 
  group_by(Species_Tmt_Pos) %>% 
  mutate(Total_Germination = cumsum(Emergence)) %>% 
  ungroup() %>% 
  mutate(Salinity = factor(Salinity, levels = c("Fresh", "25", "40"))) %>% 
  group_by(Species, Salinity, Day) %>% 
  summarize(SE = std.error(Total_Germination),
                   Total_Germination = mean(Total_Germination)) %>% 
  ungroup() %>% 
  as_tibble() %>% 
  ggplot(aes(x = Day, y = Total_Germination, color = Species)) +
  geom_point(size = 4) +
  geom_line() +
  labs(title = "Cumulative Germination",
       x = "Monitoring Day",
       y = "Mean Cumulative Germination") +
  theme(axis.title = element_text(size = 18),
        plot.title = element_text(size = 22, hjust = 0.5),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        ) +
  scale_color_manual(values = vec) +
  ylim(0, 100) + # Set 100 upper limit since that is what we seeded
  facet_wrap(~ Salinity) +
  theme_bw()
       
```

It is pretty clear from the above graph that salinity DID have an effect on species
germination, and perhaps that we would expect differing magnitudes of reduction 
between species as we move from fresh to 25 ppt salinity. 

We shouldn't have to go to crazy efforts to prove what we already know by looking 
at our raw data. Not to mention - there is danger in applying complicated statistics
when you cannot easily explain their purpose.

With that said, we will go forward with the use of this model.

Let's explore salinity effect with model estimates.
```{r sal effect, echo = FALSE}
emmip(G_Mod, Species ~ Salinity, CIs = T)
```
  These error bars are a lot more reasonable!

Other exploration graphs...

```{r explore mod graphs, echo = FALSE}
emmip(G_Mod, Salinity ~ Species, CIs = T, type = "response")

# Or without confidence intervals

emmip(G_Mod, Salinity ~ Species, CIs = F, type = "response")
```

Check for significance of fixed effects
```{r car anova, echo = FALSE}
car::Anova(G_Mod)
```
Yup, as expected: the interaction between species and salinity is significant.
We can use emmeans to view real-world (back transformed) model estimates of species
germination at each salinity level.

```{r emmeans, echo = FALSE}
emmeans(G_Mod,  ~ Salinity | Species, type = "response") 
```
Whew! We did a lot of work here. Now that we know our interaction is significant,
we can try to tease out which combination of effects are significant. However, that
will have to wait for another chapter. See you soon!