# Drawing further inference
We now have a model that shows a significant interaction between salinity 
treatments and species. We will proceed with custom contrast statements to 
pinpoint what treatments, and what species differed significantly from each other.

Let's load any required packages and our data.
```{r load packages, echo=F}
library(emmeans)
library(glmmTMB)
library(car)
library(DHARMa)
library(blme)


setwd("../Processed_Data")
load("Model_object.RData") # Load model object here
load("DF_Germination-Fresh-25-Only.RData") # Load dataframe here
```

We want to specify contrasts between the REDUCTION in germination between the
freshwater and the 25 ppt salinity treatment. To do that, we need to subtract
the model estimate of total germination of the 25 ppt salinity level from the 
freshwater salinity level. 

We can do this using the emmeans object output.

```{r emm grid}
emm1 <- emmeans(G_Mod2, ~ Salinity | Species) 
con1 <- contrast(emm1, interaction = "pairwise")
```

The emm1 object here grabs the model estimates from our two salinity levels; the
con1 object then subtracts the two estimates from each other within species for 
the two salinities (and runs a pairwise test). Our research question is: Are there
different rates of reduction BETWEEN species at different salinity levels?
We can use a pairwise comparison to answer this question.

## Pairwise comparison

```{r pairwise}
pairs(con1, by = NULL)
```
## Conclusions
Increasing salinity reduced germination for our three species, and this reduction varied
between species. Specifically, Allenrolfea occidentalis (ALOC) showed a lesser reduction
in germination at a higher salinity level than Phragmites australis. This research
shows promise for species selection when restoring invaded sites from seed. Often,
invasive species like Phragmites grow back from seed due to the high density of 
remaining seeds in the seed bank; this research shows that Allenrolfea can compete
with Phragmites germination at freshwater, and outpace Phragmites germination at a 
more saline condition. 

Other species, such as Salicornia rubra, experienced low germination regardless of
salinity. This shoud not write off using Salicornia in future restoration experiments,
but rather it leads to more questions about different dormancy breaking methods or ways
to help native playa species establish in a restoration context.
